---
title: "Final Presentation"
author: "Franciszek Wiercigroch"
subtitle: "05/12/2025"
format:
  revealjs:
    scrollable: true
    slide-number: true
    show-slide-number: all
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
editor: visual
---

## Project 1 – Objects Launched Into Space

::: panel-tabset
### Overview

-   Number of objects launched into space from 1960 until present day today
:::

::: panel-tabset
### Code Visualization

```{r}
library(ggplot2)
library(dplyr)
library(readr)

outer_space_objects <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-04-23/outer_space_objects.csv')

launches_per_year <- outer_space_objects %>%
  group_by(Year) %>%
  summarize(total_objects = sum(num_objects, na.rm = TRUE)) %>%
  arrange(Year)

max_launch_year <- launches_per_year %>%
  filter(total_objects == max(total_objects))

ggplot(launches_per_year, aes(x = Year, y = total_objects)) +
  geom_line(color = "blue") +
  geom_point(alpha = 0.2) +
  geom_text(aes(label = ifelse(total_objects == max(total_objects), 
                               paste0(total_objects, " launches"), ""))) +
  labs(title = "Trend of Space Object Launches by Year",
       x = "Year",
       y = "Number of Objects Launched") +
  theme_minimal()
```
:::

## Project 5 – Traffic Stops by Age Group

### Overview

-   Using the Stanford Open Policing Project\
-   Comparing Charlotte, Chicago, and New Orleans\
-   Focused on age-based disparities in search and arrest rates

### Search Rate by Age

-   New Orleans has the highest search rates across all age groups

-   Young drivers (under 30) face the most frequent searches

-   Charlotte has moderate search rates that decline with age

-   Chicago has consistently low search rates regardless of age

```{r, echo=FALSE}
con_traffic <- DBI::dbConnect(
  RMariaDB::MariaDB(),
  dbname = "traffic",
  host = Sys.getenv("TRAFFIC_HOST"),
  user = Sys.getenv("TRAFFIC_USER"),
  password = Sys.getenv("TRAFFIC_PWD")
)
```

```{sql, connection=con_traffic, output.var="age_data", echo=FALSE}
SELECT 
  city,
  CASE 
    WHEN subject_age < 18 THEN '<18'
    WHEN subject_age BETWEEN 18 AND 30 THEN '18–30'
    WHEN subject_age BETWEEN 31 AND 45 THEN '31–45'
    WHEN subject_age BETWEEN 46 AND 60 THEN '46–60'
    ELSE '60+'
  END AS age_group,
  COUNT(*) AS total_stops,
  ROUND(AVG(IFNULL(search_conducted, 0)), 4) AS search_rate
FROM (
  SELECT 'Charlotte' AS city, subject_age, search_conducted
  FROM nc_charlotte_2020_04_01
  WHERE subject_age IS NOT NULL
  UNION ALL
  SELECT 'Chicago' AS city, subject_age, search_conducted
  FROM il_chicago_2023_01_26
  WHERE subject_age IS NOT NULL
  UNION ALL
  SELECT 'New Orleans' AS city, subject_age, search_conducted
  FROM la_new_orleans_2020_04_01
  WHERE subject_age IS NOT NULL
) AS combined
GROUP BY city, age_group
ORDER BY city, age_group;
```

```{r}
library(ggplot2)

ggplot(age_data, aes(x = age_group, y = search_rate, fill = city)) +
  geom_col(position = "dodge") +
  labs(
    title = "Search Rate by Age Group",
    x = "Age Group",
    y = "Search Rate",
    fill = "City"
  )
```

### Arrest Rate by Age

-   Minors in New Orleans are arrested in nearly 28% of stops

-   Chicago shows high arrest rates for ages 18–45 despite low search rates

-   Arrest trends reveal major differences in how cities enforce stops by age

```{sql, connection=con_traffic, output.var="arrest_data", echo=FALSE}
SELECT 
  city,
  CASE 
    WHEN subject_age < 18 THEN '<18'
    WHEN subject_age BETWEEN 18 AND 30 THEN '18–30'
    WHEN subject_age BETWEEN 31 AND 45 THEN '31–45'
    WHEN subject_age BETWEEN 46 AND 60 THEN '46–60'
    ELSE '60+'
  END AS age_group,
  ROUND(AVG(IFNULL(arrest_made, 0)), 4) AS arrest_rate
FROM (
  SELECT 'Charlotte' AS city, subject_age, arrest_made
  FROM nc_charlotte_2020_04_01
  WHERE subject_age IS NOT NULL
  UNION ALL
  SELECT 'Chicago' AS city, subject_age, arrest_made
  FROM il_chicago_2023_01_26
  WHERE subject_age IS NOT NULL
  UNION ALL
  SELECT 'New Orleans' AS city, subject_age, arrest_made
  FROM la_new_orleans_2020_04_01
  WHERE subject_age IS NOT NULL
) AS combined
GROUP BY city, age_group
ORDER BY city, age_group;
```

```{r}
ggplot(arrest_data, aes(x = age_group, y = arrest_rate, fill = city)) +
  geom_col(position = "dodge") +
  labs(
    title = "Arrest Rate by Age Group",
    x = "Age Group",
    y = "Arrest Rate",
    fill = "City"
  )
```
